Supporting Information:\\ NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2019-01-07 Manuscript/manuscriptPSsuppl.tex

\subsection{Lipid17}
{\it DOPS and POPS bilayers.} \\
The data in Table \ref{PSsystems} and Refs \citenum{lipid17POPSjcions,lipid17POPSff99ions,lipid17DOPSjcions,lipid17DOPSff99ions}
\todo{B. Kav and M. Miettinen, please write simulation details} \\

\noindent
{\it POPC:POPS (5:1) mixtures with additional potassium and sodium ions} \\
The data in Table \ref{mixedIONsystems} and Refs \citenum{POPCpopsLIPID17withKCI,POPCpopsLIPID17withK,POPCpopsLIPID17withNaCI,POPCpopsLIPID17withNa}
\todo{B. Kav and M. Miettinen, please write simulation details} \\

>>>>

\subsection{Lipid17}
{\it DOPS and POPS bilayers.} \\

The structure of the bilayers with 128 DOPS or 128 POPS lipids were
obtained from CHARMM-GUI~\cite{lee16,jo18}. Bilayers are solvated with
4480 water molecules resulting in 31.1 water molecules per lipid. For
each system, 128 NaCl molecules are added to neutralize the system.\\

Each bilayer structure was first minimized for 2500 steps with steepest
descent and an additional 2500 step with the conjugent gradient
algorithms while restrainig the solvent and ion molecules with a force
constant 500~kcal~mol$^{-1}$~\AA{}$^{-1}$. An additional minimization
step was used with the same parameters after removing the constraints on
the solvent and ion molecules. During the minimization procedure,
non-periodic boundary conditions were applied, no pressure scaling was
used and none of the bonds were restrained. A three step heating
procedure was applied: in the first heating step the system temperature
is increased from 0.0~K to 100.0~K using Langevin thermostat with
5.0~ps$^{-1}$ collision frequency while restraining the lipid positions
with harmonic springs of force constant
20.0~kcal~mol$^{-1}$~\AA{}$^{-1}$ at constant volume without periodic
boundary conditions. In the second heating step, the system temperature
was increased from 100.0~K to 200.0~K using Langevin thermostat with
5.0~ps$^{-1}$ collision frequency within 10000 steps while restraining
the lipid positions with harmonic springs of force constant
10.0~kcal~mol$^{-1}$~\AA{}$^{-1}$ at constant volume. In the third
heating step, the system temperature is increased from 200.0~K to
303.0~K within 10000 steps using Langevin thermostat with 5.0~ps$^{-1}$
collision frequency while restraining the lipid positions with harmonic
springs of force constant 10.0~kcal~mol$^{-1}$~\AA{}$^{-1}$ at 1~bar
constant pressure using Berendsen barostat with 3.0~ps relaxation time.
From the second heating step onward, periodic boundary conditions in all
dimensions were applied. The semi-isotropic pressure coupling along
\textit{xy}-plane was employed and the surface tension was set to
0.0~dyn explicitely. After the third heating step, each structure is
equilibriated for 100~ns at 1~bar and 303~K constant pressure using
Berendsen barostat and temperature using Langevin thermostat,
respectively. For the equilibriation step, semi-isotropic pressure
coupling along \textit{xy}-plane was employed and the surface tension
was set to 0.0~dyn explicitely. For all heating and equilibration steps,
the lengths of the hydrogen containing bonds were restrained using SHAKE
algorithm and a time step of 2~fs was used. In all steps with periodic
boundary conditions, the non-bonded interactions are calculated with
particle mesh Ewald~\cite{essman95} using 1.0~nm cutoff.\\

For the production runs, the temperature and pressure were set to 303~K
and 1~bar, using Langevin thermostat with 1.0~ps$^{-1}$ collision
frequency and Berendsen barostat with 1.0~ps relaxation time,
respectively. Periodic boundary conditions were applied in all
dimensions and the non-bonded interactions are calculated with particle
mesh Ewald~\cite{essman95} using 1.0~nm cutoff. The lengths of the
hydrogen containing bonds were constrained using SHAKE algorithm and a
time step of 2~fs was used. Trajectories are saved in every 10~ps
intervals. Two independent trajectories of 500~ns long was generated and
the last 200~ns portions were used for the analysis.\\

In all simulations, TIP3P~\cite{jorgensen1983comparison} was used as the solvent. SETTLE~\cite{miyamoto92} algorithm was employed 
to constrain the bonds in TIP3P water molecules. Lipid parameters were obtained from 
Amber Lipid17 force field. For the ions, one set of simulations were run with Joung-Cheatham
ion parameters~\cite{joung2008determination} and one set of simulations were run with ff99SB~\cite{aqvist90} ion parameters. The data in Table~\ref{PSsystems} and Refs~\citenum{lipid17POPSjcions,lipid17POPSff99ions,lipid17DOPSjcions,lipid17DOPSff99ions} contain the details of the simulations and the raw trajectories, respectively.

\noindent
{\it POPC:POPS (5:1) mixtures with additional potassium and sodium ions} \\

The initial structure of the bilayers with 120 POPC and 24 POPS lipids were obtained from CHARMM-GUI~\cite{lee16,jo18}. Bilayers were solvated with 5760 TIP3P water molecules~\cite{jorgensen1983comparison}. For the simulations of the lipids and ions, the Amber Lipid 17 and ff99SB force fields were used, respectively. The simulation steps are the same as the the simulations of DOPS and POPS bilayers described above, with the only difference that the total simulation time was 300~ns, of which last 200~ns was used for the analysis. The number of ions in each specific system is given in Table~\ref{mixedIONsystems}. The trajectory files can be found in Refs~\citenum{POPCpopsLIPID17withKCI,POPCpopsLIPID17withK,POPCpopsLIPID17withNaCI,POPCpopsLIPID17withNa}.\\
